---
title: "Research Proposal"
author: 
  - name: "Jonas Loopers Davidsen"
format:
  html:
    toc: true
    toc-location: left
---
> Disclaimer: This document is merely a research proposal and therefore still subject to changes along the process. ongoing research project conducted as a part of Macroeconometrics (ECOM9007) at The University of Melbourne, Australia.\

```{r download data}
#| echo: false
#| message: false
#| warning: false
library(fredr)
library(quantmod)

fredr_set_key("2ffcd7e6c4f6e03de63ae1a03e4c3e6e")

M2 <- fredr(series_id = "M2SL", observation_start = as.Date("1987-01-01"), observation_end = as.Date("2022-12-31"))
FF <- fredr(series_id = "FEDFUNDS", observation_start = as.Date("1987-01-01"), observation_end = as.Date("2022-12-31"))
CPI <- fredr(series_id = "USACPIALLMINMEI", observation_start = as.Date("1987-01-01"), observation_end = as.Date("2022-12-31"))
HP <- fredr(series_id = "CSUSHPISA", observation_start = as.Date("1987-01-01"))
SPX <- getSymbols("^GSPC", src = "yahoo", auto.assign = FALSE, from = "1987-01-01", to = "2022-12-31")
IP <- fredr(series_id = "INDPRO", observation_start = as.Date("1987-01-01"), observation_end = as.Date("2022-12-31"))
```

## The question objective and motivation

### The objective

The objective of this research project is to investigate the effects of a money supply shock on asset prices and inflation in the US economy.

### The research question

How does money supply affect asset prices and inflation and what are the implications for monetary policy and financial stability?

### Motivation

Understanding the effects of money supply shocks on asset prices and inflation is crucial for policymakers and investors. Despite extensive Quantitative Easing (QE) programs following the financial crisis in 2008 inflation remained well under the target level in many advanced economies. Rather, the increase in the money supply primarily seemed to inflate asset prices instead of prices in general and consequently also failed to stimulate aggregate demand. However, following the Covid-19 pandemic central banks quite drastically expanded their QE programs and thereby increased the money supply to unprecedented levels in response to the economic downturn, which contributed to high inflation in many countries. This raises questions about the effectiveness of monetary policy in stimulating the economy and controlling inflation. It is therefore of interest to investigate the mechanisms through which a money supply shock affects different economic variables such as asset prices and inflation in light of economic and financial stability.

## Data and their properties

So far the intention is to include the following six variables for the US economy in the SVAR model.

* $M_t$: M2 aggregate from FRED Database

* $SPX_t$: SP500 index from Yahoo Finance
* $HP_t$: S&P/Case-Shiller U.S. National Home Price Index from FRED Database
* $CPI_t$: Consumer Price Index: All Items for the US from FRED Database
  + Motivation: Given my focus on the relationship between money supply, asset prices and inflation, a measure for those three variables are needed. As a measure for money supply the M2 aggregate is chosen as it serves as a good proxy for the availability of liquidity in the economy. As measures for asset prices, both stock prices and house prices are included. These two types of assets are big components of the total assets in the economy and provide a way to investigate the transmission mechanism of monetary policy to the real economy. Further, the CPI is chosen as it is commonly used for a measure of headline inflation.
* $ff_t$: Effective Fed Funds Rate from FRED Database
* $IP_t$:  Industrial Production: Total Index
  + Motivation: The effective fed funds rate is the rate at which banks lend and borrow funds from each other overnight and is obviously heavily influenced by the actual fed funds rate. Industrial production is a measure for monthly US real activity and is chosen since actual GDP data is only available for quarterly data. These two variables are important to include as they also play a crucial role in the relationship between money supply, asset prices and inflation. 

My sample period will be from M1 1987 - M12 2022 as data for $HP_t$ only goes back to this period and the frequency of the data is monthly.



### Transformation and visualisation of the variables

Since the effective fed funds rate, $ff_t$, is in percentages this variable is not being transformed. For the money supply, $M_t$, the log is taken:

$m_t=\log(M_t)$

The same goes for $SPX_t$, whereafter MoM returns are computed by taking first differences:

$spx_t=\log(SPX_t-SPX_{t-1})$

For $HP_t$ I compute YoY returns by taking the differences with respect to the 12th lag:

$hp_t=\log(HP_t-HP_{t-12})$

To compute YoY inflation the same is being done for $CPI_t$:

$inf_t=\log(CPI_t-CPI_{t-12})$

And last but not least the YoY growth of $IP_t$ is being computed:

$ip_t=\log(IP_t-IP_{t-12})$

We get the following plots for the variables:

```{r plotting data}
library(xts)

## Log-transform data for M2, CPI & HP and transform SPX to monthly data

M2 <- xts::xts(log(M2$value), M2$date)
CPI <- xts::xts(log(CPI$value), CPI$date)
HP <- xts::xts(log(HP$value), HP$date)
FF <- xts::xts(FF$value, FF$date)
IP <- xts::xts(IP$value, IP$date)

SPX <- xts::xts(log(SPX[,6]), order.by=as.Date(index(SPX)))
SPX <- xts::apply.monthly(SPX, mean)

#Take first differences for HP, SPX and IP 

HP <- diff(HP, lag = 12, na.pad = TRUE) # YoY return
SPX <- diff(SPX) #MoM return
INF <- 100*diff(CPI, lag = 12, na.pad = TRUE)

#Remove first 12 observations from all variables to ensure same dimensions

INF <- INF[-c(1:12)]
M2 <- M2[-c(1:12)]
FF <- FF[-c(1:12)]
SPX <- SPX[-c(1:12)]
HP <- HP[-c(1:12)]
IP <- IP[-c(1:12)]

index(SPX) <- as.Date(format(index(SPX), "%Y-%m-01")) #Make the dates in all series align

#Plotting the time series

par(mfrow=c(3,2), mar=c(4,4.5,2,2),cex.axis=1.5, cex.lab=1.5)
plot.ts(M2, main = "", ylab = expression(m[t]), xlab = "time", col = "magenta", lwd = 2)
plot.ts(SPX,main = "", ylab=expression(spx[t]), xlab = "time", col="magenta",lwd=2)
plot.ts(HP,main = "", ylab=expression(hp[t]), xlab = "time", col="magenta",lwd=2)
plot.ts(INF,main = "", ylab=expression(inf[t]), xlab = "time", col="magenta",lwd=2)
plot.ts(FF,main = "", ylab=expression(ff[t]), xlab = "time", col="magenta",lwd=2)
plot.ts(IP,main = "", ylab=expression(ip[t]), xlab = "time", col="magenta",lwd=2)
```
## The model and hypothesis
 
$[m_t$   $spx_t$   $hp_t$   $inf_t$   $ff_t$   $ip_t]'$


\begin{align}
B_0y_t &= b_0 + B_1y_{t-1}+\dots+B_py_{t-p}+u_t\\
u_t|Y_{t-1} &\sim \; iid(\textbf{0}_N,I_N)
\end{align}












